RIScatter: Unifying Backscatter Communications and Reconfigurable Intelligent Surface From a Probabilistic Perspective

### 2022-11-03 main.tex
:dizzy: Manuscript 221103

	\gls{bc} nodes harvest energy from and modulate information over an external radio wave by manipulating the magnitude, phase, and/or frequency of the scattered signal.

@@

	\gls{ris} is a programmable reflector array that adapts the phase shift response to enhance or suppress signal strength in specific directions.
	In this paper, we show how those three seemingly different technologies can be unified to leverage their benefits simultaneously into a single architecture called RIScatter.
	RIScatter is a new paradigm for future wireless networks and consists of multiple dispersed or co-located passive scatter nodes, whose reflection states can be adapted to partially engineer the wireless channel and partially modulate information onto the scattered wave.
	This contrasts with \gls{bc}/\gls{sr} (resp. \gls{ris}) where states are exclusively a function of information symbols (resp. \gls{csi}).

@@

	This enables RISscatter to softly bridge, generalize, and outperform \gls{bc}, \gls{sr}, \gls{ris}; boil down to any of those under specific reflection states; or evolve in a mixed form for universal hardware design and heterogeneous traffic control.
	For a typical setup, we characterize the achievable primary-(total-)backscatter rate region by optimizing input distribution at the nodes, active beamforming at the \gls{ap}, and backscatter detector at the user.
	Simulation results demonstrate RIScatter nodes can exploit the additional propagation paths to smoothly transition between backscatter modulation and passive beamforming via smart input distribution design, and the proposed receiver significantly reduces the decoding complexity and effectively accommodates the double modulation and signal difference in active-passive coexisting network.

@@

		Those \gls{ris}-empowered \gls{bc}/\gls{sr} designs involve advanced hardware architecture and high optimization complexity.

@@

	\begin{subsection}{Contributions}
		As presented in Fig. \subref*{fg:riscatter}, we propose RIScatter as a novel scattering protocol that generalizes \gls{bc}, \gls{sr}, and \gls{ris}.
		The contributions of this paper are summarized as follows.

		\emph{First,} we propose RIScatter nodes that adapt the reflection state distribution of passive scatter devices based on \gls{csi} and link weights, in order to flexibly transition between backscatter modulation and passive beamforming.
		Scattering sources of \gls{bc}/\gls{ambc}/\gls{sr} and reflecting elements of \gls{ris} can be regarded as special cases with uniform and degenerate input distributions, respectively.
		On the contrary, adaptive RIScatter encoding boils down to deterministic \gls{ris} when primary link is prioritized, and achieves higher backscatter rate than conventional line coding when backscatter link is prioritized.
		Multiple RIScatter nodes can be either co-located to enable joint distribution design for improved total backscatter rate, or dispersed to guarantee uniformly good performance for both links.

@@

\begin{section}{Scattering Principles}
	\gls{rf} wave scattering and reflecting are usually realized by a \emph{variable-load antenna} or \emph{programmable metamaterial} and described by a unified signal model \cite{Liang2022}.
	A typical antenna-based scatterer consists of an integrated antenna, a load-switching modulator, an energy harvester, and on-chip components (e.g., microcontroller and sensors) \cite{Dobkin2012}.
	It first receives the impinging signals, then reradiates some back to the space and dissipates the remaining.
	In comparison, a typical metamaterial-based scatterer comprises an outer metamaterial layer of numerous sub-wavelength metallic/dielectric patches with tunable permittivity/permeability, a middle copper plate layer that reflects residual to avoid leakage, an inner circuit board layer that adjusts the amplitude and phase responses of patches, and an integrated microcontroller/\glsentryshort{fpga} that coordinates with the network and controls the circuit \cite{Wu2020}.
	Ideally, it reflects the incident waves at the air-metamaterial boundary without receiving them, and mainly applies a phase shift on the reflected wave.
	In practice, both types of scatterers have finite reflection states with non-zero reflection loss, and the scattered signal can be decomposed into the \emph{structural} and \emph{antenna} mode components \cite{Hansen1989}.
	The former consistently contributes to environment multipath and can be modelled within \gls{csi}, while the latter depends on impedance mismatch and can be used for backscatter modulation and/or passive beamforming.
	For a scatter node with $M$ reflection states, the reflection coefficient at state $m \in \mathcal{M} \triangleq \{1,\ldots,M\}$ is
	\begin{equation}
		\Gamma_m = \frac{Z_m - Z^*}{Z_m + Z},
		\label{eq:reflection_coefficient}
	\end{equation}
	where $Z_m$ is the antenna load (resp. metamaterial unit) impedance at state $m$ and $Z$ is the antenna input (resp. medium characteristic) impedance.
	% Backscatter modulation and passive beamforming are realized by manipulating the reflection coefficient.
	\gls{bc}, \gls{ambc} and \gls{sr} employ scatter nodes as information sources that encode message by \emph{probabilistically switching} between different states.
	For $M$-ary \gls{qam}, constellation point $c_m$ maps to reflection coefficient $\Gamma_m$ by \cite{Thomas2012a}
	\begin{equation}
		\Gamma_m = \alpha \frac{c_m}{\max_{m'} \lvert c_{m'} \rvert},
		\label{eq:backscatter_modulation}
	\end{equation}
	where $\alpha \in \mathbb{I}$ is the amplitude scattering ratio at the direction of interest.
	In contrast, \gls{ris} employ reflecting elements as channel reconfigurators that enable constructive/destructive signal superposition by \emph{deterministically choosing} the reflection pattern.
	For a \gls{ris} element with $M$ available states, phase shift $\theta_m$ maps to reflection coefficient $\Gamma_m$ by \cite{Wu2018}
	\begin{equation}
		\Gamma_m = \beta_m \exp(j \theta_m),
		\label{eq:passive_beamforming}
	\end{equation}
	where $\beta_m \in \mathbb{I}$ is the amplitude scattering ratio of state $m$.\footnote{Most existing \gls{ris} literatures assume lossless reflection $\beta_m=1$, $\forall m$.}
\end{section}

\begin{section}{RIScatter}
	\begin{subsection}{Concepts}
		As a generalization of \gls{bc}, \gls{sr} and \gls{ris}, RIScatter is a passive scattering protocol that coexists with an active primary system in a flexible and mutualistic manner.
		RIScatter nodes leverage \gls{csi}- and weight-based input distribution design to smoothly transition between backscatter modulation and passive beamforming.
		It can be implemented by adding an integrated receiver \cite{Kim2021a} and adaptive encoder \cite{He2020e} to off-the-shelf passive backscatter tags.
		The block diagram, equivalent circuit, and scatter model of a RIScatter node are illustrated in Fig. \ref{fg:riscatter_node}, while the input distribution and time structure comparison of scattering applications are shown in Fig. \ref{fg:scatter_comparison}.
		\begin{figure*}[!t]
			\centering
			\subfloat[Block Diagram]{
				\resizebox{0.32\linewidth}{!}{
					\input{assets/illustration/block_diagram.tex}
				}
				\label{fg:block_diagram}
			}
			\subfloat[Equivalent Circuit]{
				\resizebox{0.37\linewidth}{!}{
					\input{assets/illustration/equivalent_circuit.tex}
				}
				\label{fg:equivalent_circuit}
			}
			\subfloat[Scatter Model]{
				\resizebox{0.28\linewidth}{!}{
					\input{assets/illustration/scatter_model.tex}
				}
				\label{fg:scatter_model}
			}
			\caption{
			Block diagram, equivalent circuit, and scatter model of a RIScatter node.
			The solid and dashed vectors represent signal and energy flows.
			The scattering antenna behaves as a constant power source, where the voltage $V_0$ and current $I_0$ are introduced by incident electric field $\vec{E}_{\mathrm{I}}$ and magnetic field $\vec{H}_{\mathrm{I}}$ \cite{Huang2021}.
			}
			\label{fg:riscatter_node}
		\end{figure*}
		\begin{figure}[!t]
			\centering
			\subfloat[Input Distribution]{
				\resizebox{0.8\columnwidth}{!}{
					\input{assets/illustration/input_distribution.tex}
				}
				\label{fg:input_distribution}
			}
			\\
			\subfloat[Time Structure]{
				\resizebox{0.8\columnwidth}{!}{
					\input{assets/illustration/time_structure.tex}
				}
				\label{fg:time_structure}
			}
			\caption{
				Input distribution and time structure of \gls{bc}, \gls{ambc}, \gls{sr}, \gls{ris}, and RIScatter.
				``PB'' means primary block and ``BB'' means backscatter block.
			}
			\label{fg:scatter_comparison}
		\end{figure}
		Instead of using fully random or fully deterministic reflection pattern over time, each RIScatter node semi-randomly chooses the reflection state for each backscatter block with guidance of input probability distribution $P(\Gamma_m)$ at state $m$.
		Such an adaptive backscatter coding boils down to degenerate distribution of \gls{ris} when primary link is prioritized, and outperforms conventional line coding with equiprobable inputs (e.g., FM0 for \gls{rfid}) when backscatter link is prioritized.
		Besides, joint encoding over multiple co-located RIScatter nodes can further boost the total backscatter rate.
		Relevant \gls{csi} can be acquired by existing low-power estimation techniques.
		For dispersed RIScatter nodes, the cascaded transmitter-node-receiver channels can be estimated by sequential \cite{Bharadia2015,Yang2015b,Guo2019g} or parallel approaches \cite{Jin2021a} originally proposed for \gls{bc}.
		For co-located RIScatter nodes, the estimation can be simplified by group-based \cite{Zheng2019} and hierarchical \cite{You2019} trainings originally proposed for \gls{ris}.

@@

	Problem \eqref{op:weighted_sum_rate} generalizes conventional \gls{bc}/\gls{sr} by allowing \gls{csi}- and weight-adaptive input distribution and detection region design.
	It also generalizes the discrete \gls{ris} phase shift selection by allowing stochastic reflection (i.e., relaxing the feasible domain from the vertices of $M$-dimensional probability simplex to the simplex itself).
	Since problem \eqref{op:weighted_sum_rate} is highly non-convex, we propose a \gls{bcd} algorithm that iteratively updates $\{\boldsymbol{p}_k\}_{k \in \mathcal{K}}$, $\boldsymbol{w}$ and $\boldsymbol{t}$ until convergence.

@@

\begin{section}{Conclusion}
	This paper introduced RIScatter as a low-power scattering protocol that unifies backscatter modulation and dymanic passive beamforming by smart input distribution and practical receiver design.
	Starting from scattering principles, we showed how RIScatter nodes include scattering source of \gls{bc}/\gls{sr} and reflecting element of \gls{ris} as special cases, how they can be built over existing passive scatter devices, and how they simultaneously encode self information and assist legacy transmission.
	We also propose a low-complexity RIScatter receiver that preserves the benefits of backscatter modulation and passive beamforming.
	The achievable primary-total-backscatter rate region is then studied for a single-user multi-node RIScatter system, where the input distribution, active beamforming, and decision thresholds are iteratively updated.
	Numerical results not only demonstrated the effectiveness of the proposed algorithms, but also emphasized the importance of adaptive input distribution and cooperative decoding on both primary and backscatter subsystems.

>>>>

	% \gls{bc} nodes harvest energy from and modulate information over an external radio wave by manipulating the magnitude, phase, and/or frequency of the scattered signal.
	% \gls{bc} nodes harvest energy from and modulate information over an external electromagnetic wave by manipulating the magnitude, phase, and/or frequency of the scattered signal.
	\gls{bc} nodes harvest energy from and modulate information over an external electromagnetic wave.

@@

	% \gls{ris} is a programmable reflector array that adapts the phase shift response to enhance or suppress signal strength in specific directions.
	\gls{ris} adapts its phase shift response to enhance or suppress signal strength in specific directions.
	In this paper, we show how those two seemingly different technologies (and their derivatives) can be unified to leverage their benefits simultaneously into a single architecture called RIScatter.
	As a new paradigm for future wireless networks, RIScatter consists of multiple dispersed or co-located passive scatter nodes whose reflection states can be adapted to partially engineer the wireless channel and partially modulate information onto the scattered wave.
	This contrasts with \gls{bc} (resp. \gls{ris}) where states are exclusively a function of information symbols (resp. \gls{csi}).

@@

	This enables RISscatter to softly bridge, generalize, and outperform \gls{bc} and \gls{ris}; boil down to any of those under specific reflection states; or evolve in a mixed form for universal hardware design and heterogeneous traffic control.
	For a single-user multi-node RIScatter network, we characterize the achievable primary-(total-)backscatter rate region by optimizing input distribution at the nodes, active beamforming at the \gls{ap}, and backscatter detection region at the user.
	Simulation results demonstrate RIScatter nodes can exploit the additional propagation paths to smoothly transition between backscatter modulation and passive beamforming via smart input distribution design, and the proposed practical receiver effectively accommodates the double modulation and signal difference in active-passive coexisting networks.

@@

		Those \gls{ris}-empowered \gls{bc} designs involve advanced hardware architecture and high optimization complexity.

@@

	\begin{subsection}{Contributions}
		As presented in Fig. \subref*{fg:riscatter}, we propose RIScatter as a novel scattering protocol that generalizes \gls{bc} and \gls{ris}.
		The contributions of this paper are summarized as follows.

		\emph{First,} we propose RIScatter nodes that adapt the reflection state distribution of passive scatter devices based on \gls{csi} and link weights, in order to flexibly transition between backscatter modulation and passive beamforming.
		Scattering sources of \gls{bc} and reflecting elements of \gls{ris} can be regarded as special cases with uniform and degenerate input distributions, respectively.
		On the contrary, adaptive RIScatter encoding boils down to deterministic \gls{ris} when primary link is prioritized, and achieves higher backscatter rate than conventional line coding when backscatter link is prioritized.
		Multiple RIScatter nodes can be either co-located to enable joint distribution design for improved total backscatter rate, or dispersed to guarantee uniformly good performance for both links.

@@

\begin{section}{Scattering Principles}
	\gls{rf} wave scattering and reflecting are usually realized by a \emph{variable-load antenna} or \emph{programmable metamaterial} and described by a unified signal model \cite{Liang2022}.
	A typical antenna-based scatterer consists of an integrated antenna, a load-switching modulator, an energy harvester, and on-chip components (e.g., microcontroller and sensors) \cite{Dobkin2012}.
	It first receives the impinging signals, then reradiates some back to the space and dissipates the remaining.
	In comparison, a typical metamaterial-based scatterer comprises an outer metamaterial layer of numerous sub-wavelength metallic/dielectric patches with tunable permittivity/permeability, a middle copper plate layer that reflects residual to avoid leakage, an inner circuit board layer that adjusts the amplitude and phase responses of patches, and an integrated microcontroller/\glsentryshort{fpga} that coordinates with the network and controls the circuit \cite{Wu2020}.
	Ideally, it reflects the incident waves at the air-metamaterial boundary without receiving them, and mainly applies a phase shift on the reflected wave.
	In practice, both types of scatterers have finite reflection states with non-zero reflection loss, and the scattered signal can be decomposed into the \emph{structural} and \emph{antenna} mode components \cite{Hansen1989}.
	The former consistently contributes to environment multipath and can be modelled within \gls{csi}, while the latter depends on impedance mismatch and can be used for backscatter modulation and/or passive beamforming.
	For a scatter node with $M$ reflection states, the reflection coefficient at state $m \in \mathcal{M} \triangleq \{1,\ldots,M\}$ is
	\begin{equation}
		\Gamma_m = \frac{Z_m - Z^*}{Z_m + Z},
		\label{eq:reflection_coefficient}
	\end{equation}
	where $Z_m$ is the antenna load (resp. metamaterial unit) impedance at state $m$ and $Z$ is the antenna input (resp. medium characteristic) impedance.
	% Backscatter modulation and passive beamforming are realized by manipulating the reflection coefficient.
	\gls{bc} employ scatter nodes as information sources that encode message by \emph{probabilistically switching} between different states.
	For $M$-ary \gls{qam}, constellation point $c_m$ maps to reflection coefficient $\Gamma_m$ by \cite{Thomas2012a}
	\begin{equation}
		\Gamma_m = \alpha \frac{c_m}{\max_{m'} \lvert c_{m'} \rvert},
		\label{eq:backscatter_modulation}
	\end{equation}
	where $\alpha \in \mathbb{I}$ is the amplitude scattering ratio at the direction of interest.
	In contrast, \gls{ris} employ reflecting elements as channel reconfigurators that enable constructive/destructive signal superposition by \emph{deterministically choosing} the reflection pattern.
	For a \gls{ris} element with $M$ available states, phase shift $\theta_m$ maps to reflection coefficient $\Gamma_m$ by \cite{Wu2018}
	\begin{equation}
		\Gamma_m = \beta_m \exp(j \theta_m),
		\label{eq:passive_beamforming}
	\end{equation}
	where $\beta_m \in \mathbb{I}$ is the amplitude scattering ratio of state $m$.\footnote{Most existing \gls{ris} literatures assume lossless reflection $\beta_m=1$, $\forall m$.}
\end{section}

\begin{section}{RIScatter}
	\begin{subsection}{Concepts}
		As a generalization of \gls{bc} and \gls{ris}, RIScatter is a passive scattering protocol that coexists with an active primary system in a flexible and mutualistic manner.
		RIScatter nodes leverage \gls{csi}- and weight-based input distribution design to smoothly transition between backscatter modulation and passive beamforming.
		It can be implemented by adding an integrated receiver \cite{Kim2021a} and adaptive encoder \cite{He2020e} to off-the-shelf passive backscatter tags.
		The block diagram, equivalent circuit, and scatter model of a RIScatter node are illustrated in Fig. \ref{fg:riscatter_node}, while the input distribution and time structure comparison of scattering applications are shown in Fig. \ref{fg:scatter_comparison}.
		\begin{figure*}[!t]
			\centering
			\subfloat[Block Diagram]{
				\resizebox{0.32\linewidth}{!}{
					\input{assets/illustration/block_diagram.tex}
				}
				\label{fg:block_diagram}
			}
			\subfloat[Equivalent Circuit]{
				\resizebox{0.37\linewidth}{!}{
					\input{assets/illustration/equivalent_circuit.tex}
				}
				\label{fg:equivalent_circuit}
			}
			\subfloat[Scatter Model]{
				\resizebox{0.28\linewidth}{!}{
					\input{assets/illustration/scatter_model.tex}
				}
				\label{fg:scatter_model}
			}
			\caption{
			Block diagram, equivalent circuit, and scatter model of a RIScatter node.
			The solid and dashed vectors represent signal and energy flows.
			The scattering antenna behaves as a constant power source, where the voltage $V_0$ and current $I_0$ are introduced by incident electric field $\vec{E}_{\mathrm{I}}$ and magnetic field $\vec{H}_{\mathrm{I}}$ \cite{Huang2021}.
			}
			\label{fg:riscatter_node}
		\end{figure*}
		\begin{figure}[!t]
			\centering
			\subfloat[Input Distribution]{
				\resizebox{0.8\columnwidth}{!}{
					\input{assets/illustration/input_distribution.tex}
				}
				\label{fg:input_distribution}
			}
			\\
			\subfloat[Time Structure]{
				\resizebox{0.8\columnwidth}{!}{
					\input{assets/illustration/time_structure.tex}
				}
				\label{fg:time_structure}
			}
			\caption{
				Input distribution and time structure of \gls{bc}, \gls{ris}, and RIScatter.
				``PB'' means primary block and ``BB'' means backscatter block.
			}
			\label{fg:scatter_comparison}
		\end{figure}
		Instead of using fully random or fully deterministic reflection pattern over time, each RIScatter node semi-randomly chooses the reflection state for each backscatter block with guidance of input probability distribution $P(\Gamma_m)$ at state $m$.
		Such an adaptive backscatter coding boils down to degenerate distribution of \gls{ris} when primary link is prioritized, and outperforms conventional line coding with equiprobable inputs (e.g., FM0 for \gls{rfid}) when backscatter link is prioritized.
		Besides, joint encoding over multiple co-located RIScatter nodes can further boost the total backscatter rate.
		Relevant \gls{csi} can be acquired by existing low-power estimation techniques.
		For dispersed RIScatter nodes, the cascaded transmitter-node-receiver channels can be estimated by sequential \cite{Bharadia2015,Yang2015b,Guo2019g} or parallel approaches \cite{Jin2021a} originally proposed for \gls{bc}.
		For co-located RIScatter nodes, the estimation can be simplified by group-based \cite{Zheng2019} and hierarchical \cite{You2019} trainings originally proposed for \gls{ris}.

@@

	Problem \eqref{op:weighted_sum_rate} generalizes conventional \gls{bc} by allowing \gls{csi}- and weight-adaptive input distribution and detection region design.
	It also generalizes the discrete \gls{ris} phase shift selection by allowing stochastic reflection (i.e., relaxing the feasible domain from the vertices of $M$-dimensional probability simplex to the simplex itself).
	Since problem \eqref{op:weighted_sum_rate} is highly non-convex, we propose a \gls{bcd} algorithm that iteratively updates $\{\boldsymbol{p}_k\}_{k \in \mathcal{K}}$, $\boldsymbol{w}$ and $\boldsymbol{t}$ until convergence.

@@

\begin{section}{Conclusion}
	This paper introduced RIScatter as a low-power scattering protocol that unifies backscatter modulation and dymanic passive beamforming by smart input distribution and practical receiver design.
	Starting from scattering principles, we showed how RIScatter nodes include scattering source of \gls{bc} and reflecting element of \gls{ris} as special cases, how they can be built over existing passive scatter devices, and how they simultaneously encode self information and assist legacy transmission.
	We also propose a low-complexity RIScatter receiver that preserves the benefits of backscatter modulation and passive beamforming.
	The achievable primary-total-backscatter rate region is then studied for a single-user multi-node RIScatter system, where the input distribution, active beamforming, and decision thresholds are iteratively updated.
	Numerical results not only demonstrated the effectiveness of the proposed algorithms, but also emphasized the importance of adaptive input distribution and cooperative decoding on both primary and backscatter subsystems.
